---
title: "ARIMA Modeling"
author: "Mac Faldet"
date: "March 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(astsa)
library(xts)
```

```{r}
# View a detailed description of AirPassengers
help(AirPassengers)
## Monthly Airline Passenger Numbers 1949-1960

# Plot AirPassengers
plot(AirPassengers)

# Plot the DJIA daily closings
plot(djia$Close)

# Plot the Southern Oscillation Index (soi) and inspect it for trend, seasonality, and homoscedasticity.
plot(soi)
```

As you can see, the AirPassengers dataset contains monthly information on airline passengers from 1949 through 1960. Note that when you plot ts objects using plot(), the data will automatically be plotted over time.

#Stationarity: stable - the mean is constant over time & correlation is constant over time
#NonStationarity: unstable 

##Differencing
As seen in the video, when a time series is trend stationary, it will have stationary behavior around a trend. A simple example is Yt=α+βt+Xt where Xt is stationary.

A different type of model for trend is random walk, which has the form Xt=Xt−1+Wt, where Wt is white noise. It is called a random walk because at time t the process is where it was at time t−1 plus a completely random movement. For a random walk with drift, a constant is added to the model and will cause the random walk to drift in the direction (positive or negative) of the drift.

```{r}
plot(diff(AirPassengers))
plot(diff(djia$Close))
```

simple differencing can remove the trend and coerce the data to stationarity. Differencing looks at the difference between the value of a time series at a certain point in time and its preceding value. That is, Xt−Xt−1 is computed.

To check that it works, you will difference each generated time series and plot the detrended series. If a time series is in x, then diff(x) will have the detrended series obtained by differencing the data. To plot the detrended series, simply use plot(diff(x)).

## Detrending Data
As you have seen in the previous exercise, differencing is generally good for removing trend from time series data. Recall that differencing looks at the difference between the value of a time series at a certain point in time and its preceding value.

In this exercise, you will use differencing diff() to detrend and plot real time series data.

```{r}
# Plot globtemp and detrended globtemp
par(mfrow = c(2,1))
plot(globtemp) 
plot(diff(globtemp))

# Plot cmort and detrended cmort
par(mfrow = c(2,1))
plot(cmort)
plot(diff(cmort))
```

## Dealing with Trend and Heteroscedasticity
Here, we will coerce nonstationary data to stationarity by calculating the return or growth rate as follows.

Often time series are generated as
  `Xt=(1+pt)Xt−1`
meaning that the value of the time series observed at time t equals the value observed at time t−1 and a small percent change pt at time t.

A simple deterministic example is putting money into a bank with a fixed interest p. In this case, Xt is the value of the account at time period t with an initial deposit of X0.

Typically, pt is referred to as the return or growth rate of a time series, and this process is often stable.

For reasons that are outside the scope of this course, it can be shown that the growth rate pt can be approximated by
  `Yt=logXt−logXt−1≈pt`
In R, pt is often calculated as diff(log(x)) and plotting it can be done in one line plot(diff(log(x))).

```{r}
# Plot GNP series (gnp) and its growth rate
par(mfrow = c(2,1))
plot(gnp)
plot(diff(log(gnp)))

# Plot DJIA closings (djia$Close) and its returns
par(mfrow = c(2,1))
plot(djia$Close)
plot(diff(log(djia$Close)))
```

## Simulating ARMA Models
As we saw in the video, any stationary time series can be written as a linear combination of white noise. In addition, any ARMA model has this form, so it is a good choice for modeling stationary time series.

R provides a simple function called `arima.sim()` to generate data from an ARMA model. For example, the syntax for generating 100 observations from an MA(1) with parameter .9 is 
`arima.sim(model = list(order = c(0, 0, 1), ma = .9 ), n = 100)`. You can also use `order = c(0, 0, 0)` to generate white noise.

In this exercise, you will generate data from various ARMA models. For each command, generate 200 observations and plot the result.

```{r}
# Generate and plot white noise
WN <- arima.sim(model = list(order = c(0, 0, 0)), n = 200)
plot(WN)

# Generate and plot an MA(1) with parameter .9 by filtering the noise
MA <- arima.sim(model = list(order = c(0, 0, 1), ma = .9), n = 200)  
plot(MA)

# Generate and plot an AR(1) with parameters 1.5 and -.75
AR <- arima.sim(model = list(order = c(2, 0, 0), ar = c(1.5, -.75)), n = 200) 
plot(AR)
```

```{r}
x <- arima.sim(list(order = c(1, 0, 0), ar = -.7), n = 200)
y <- arima.sim(list(order = c(0, 0, 1), ma = -.7), n = 200)

par(mfrow = c(1,2))
plot(x, main = "AR(1)")
plot(y, main = "MA(1)")
```

```{r}

```






```{r}
fit <- arima(log10(AirPassengers), c(0, 1, 1),
              seasonal = list(order = c(0, 1, 1), period = 12))
update(fit, method = "CSS")
update(fit, x = window(log10(AirPassengers), start = 1954))
pred <- predict(fit, n.ahead = 24)
tl <- pred$pred - 1.96 * pred$se
tu <- pred$pred + 1.96 * pred$se
ts.plot(AirPassengers, 10^tl, 10^tu, log = "y", lty = c(1, 2, 2))
```

```{r}
## full ML fit is the same if the series is reversed, CSS fit is not
ap0 <- rev(log10(AirPassengers))
attributes(ap0) <- attributes(AirPassengers)
arima(ap0, c(0, 1, 1), seasonal = list(order = c(0, 1, 1), period = 12))
arima(ap0, c(0, 1, 1), seasonal = list(order = c(0, 1, 1), period = 12),
      method = "CSS")

## Structural Time Series
ap <- log10(AirPassengers) - 2
(fit <- StructTS(ap, type = "BSM"))
par(mfrow = c(1, 2))
plot(cbind(ap, fitted(fit)), plot.type = "single")
plot(cbind(ap, tsSmooth(fit)), plot.type = "single")
```
